Precise Data Flow Analysis\\in the Presence of Correlated Method Calls

### 2014-12-16 content/appendix.tex
repeating lemmas in appendix

\subsection*{Representation of Micro Functions}
\otodo{Repeat the lemmas here before each proof.}
We start by presenting the proofs to the lemmas about the representation of micro functions.

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:equalEq}}]
    OL: Can't we just say the following?
    \begin{align*}
        \denote{\setpair {I}{U}}\\
       & = \lambda m. \lambda r. (m(r) \cap I(r))\cup U(r))\\
       & = \lambda m. \lambda r. (m(r) \cap I'(r))\cup U'(r))\\
       & = 
        \denote{\setpair {I'}{U'}}\\
    \end{align*}

@@

\otodo{The (m)(r) does not seem necessary. Can we replace $r'$ with $r$ and $m'$ with $m$?}
\begin{proof}[\textbf{Proof of Lemma~\ref{lem:equalComp}}]
For all $r\in R$ and $m\in R\to2^T$,
\begin{align*}
  &\denote{\setpair IU\circ\setpair{I'}{U'}}(m)(r)\\
  &=\denote{\setpair{\lambda r'\,.\,(I(r')\cap I'(r'))\cup U(r')}
                    {\lambda r'\,.\,(I(r')\cap U'(r'))\cup U(r')}}(m)(r)\\
  &=(m(r)\cap(I(r)\cap I'(r))\cup U(r))\cup(I(r)\cap U'(r))\cup U(r)\\
  &=(m(r)\cap I'(r))\cup I(r)\cup (U'(r)\cap I(r))\cup U(r)\\
  &=(((m(r)\cap I'(r))\cup U'(r))\cap I(r))\cup U(r)\\
  &=\left(\lambda r'.(((m(r')\cap I'(r'))\cup U'(r'))\cap I(r'))\cup U(r')\right)(r)\\
  &=\left(\lambda m'.\lambda r'.(m'(r')\cap I(r'))\cup U(r')\right)
    \left(\left(\lambda m'.\lambda r'.(m'(r')\cap I'(r'))\cup U'(r')\right)  (m)\right)(r)\\
  &=\left(\denote{\setpair IU}\circ\denote{\setpair{I'}{U'}}\right)(m)(r).\qedhere
\end{align*}
\end{proof}

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:equalMeet}}]
  For all $r\in R$ and $m\in R\to2^T$,
\begin{align*}
  \left(\denote{\setpair{I}{U} \sqcap \setpair{I'}{U'}}\right)&(m)(r)\\
  &=(\denote{\setpair{\lambda r\,.\,I(r) \cup I'(r)}{\lambda r\,.\,U(r) \cup U'(r)}})(m)(r)\\
  &=(m(r)\cap(I(r)\cup I'(r))\cup U(r)\cup U'(r)\\
  &=(m(r)\cap I(r))\cup U(r)\cup(m(r)\cap I'(r))\cup U'(r)\\
  &=\denote{\setpair{I}{U}}(m)(r)\cup\denote{\setpair{I'}{U'}}(m)(r)\\
  &=\left(\denote{\setpair{I}{U}} \sqcap \denote{\setpair{I'}{U'}}\right)(m)(r).\qedhere
\end{align*}
\end{proof}

@@

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:subsetifds}}]
    \otodo{It's Section 3.2.}
  Let $P$ be an IFDS problem. Recall from Section~\ref{sec:ideToIfds} that the result of an IFDS analysis $\resultIFDS(P)$ maps supergraph
  nodes $n\in N^*$ to sets of data-flow facts $\delta\in2^D$. Specifically,
  \begin{equation}
    \resultIFDS(P)=\lambda n\,.\,\mvp F(n)\,.
  \end{equation}
  \otodo{Where do we mention $\backEq$ in this lemma?}
  The transformation $\backCC$ is the same as $\backEq$, except that it can remove data-flow facts from the result:
  \begin{align*}
    \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
    &=\left(\lambda n'\,.\,\left\{d\,|\,\forall r\in R\,.\,\resultIDE(\transCC_R(P))(n')(d)(r)\ne\top_T\right\}\right)(n)\\
    &=\left\{d\,|\,\forall r\in R\,.\,\resultIDE(\transCC_R(P))(n')(d)(r)\ne\top_T\right\}\\
    &\subseteq\left\{d\,|\,\resultIDE(\transCC_R(P))(n)(d)\ne\topcc\right\}\\
    &=\backEq\left(\resultIDE(\transCC_R(P))(n)\right).\qedhere
  \end{align*}
\end{proof}

@@

\begin{proof}[\textbf{Proof of Lemma~\ref{lem:sound}}] 
According to~\eqref{eq:ideresult}, we can rewrite~\eqref{eq:transCcDef} as
\begin{align*}\label{eq:transCcResultCc}
  \backCC(\resultIDE(\transCC_R(P)))(n)
  &=\{d'\,|\,\forall r\in R\,.\,\mvp{\textsf{Env}}(n,\,d'))(r)\ne\top_T\}\\
  &=\left\{d'\,|\,\forall r\in R\,.\bigsqcap_{q\in\textsf{VP}(n)}\menv(q)(\Omega)(d')(r)\ne\top_T\right\}\\
  &=\left\{d'\,|\,\forall r\in R\,.\bigsqcap_{q\in\textsf{VP}(n)}\mpddef q {d'}(r)\ne\top_T\right\}.
\end{align*}
  According to~Lemma~\ref{lem:sound3}, since $d\in M_F(p)(\varnothing)$, then for any $r\in R$, $\mpd(r)\ne\top_T$. 
  Since $\mpd(r)$ is a non-empty set that is contained in $\bigsqcap_{q\in\textsf{VP}(n)}\mpddef q {d}(r)$, it follows that $$\bigsqcap_{q\in\textsf{VP}(n)}\mpddef q {d}(r)\ne\top_T.$$ Therefore, $d\in \backCC(\resultIDE(\transCC_R(P)))(n)$.
\end{proof}

>>>>

\subsection*{Representation of Micro Functions}
We start by presenting the proofs to the lemmas about the representation of micro functions.

\equalEq*
\begin{proof}
    OL: Can't we just say the following?
    \begin{align*}
        \denote{\setpair {I}{U}}\\
       & = \lambda m. \lambda r. (m(r) \cap I(r))\cup U(r))\\
       & = \lambda m. \lambda r. (m(r) \cap I'(r))\cup U'(r))\\
       & = 
        \denote{\setpair {I'}{U'}}\\
    \end{align*}

@@

\otodo{The (m)(r) does not seem necessary. Can we replace $r'$ with $r$ and $m'$ with $m$?}
\equalComp*
\begin{proof}
For all $r\in R$ and $m\in R\to2^T$,
\begin{align*}
  &\denote{\setpair IU\circ\setpair{I'}{U'}}(m)(r)\\
  &=\denote{\setpair{\lambda r'\,.\,(I(r')\cap I'(r'))\cup U(r')}
                    {\lambda r'\,.\,(I(r')\cap U'(r'))\cup U(r')}}(m)(r)\\
  &=(m(r)\cap(I(r)\cap I'(r))\cup U(r))\cup(I(r)\cap U'(r))\cup U(r)\\
  &=(m(r)\cap I'(r))\cup I(r)\cup (U'(r)\cap I(r))\cup U(r)\\
  &=(((m(r)\cap I'(r))\cup U'(r))\cap I(r))\cup U(r)\\
  &=\left(\lambda r'.(((m(r')\cap I'(r'))\cup U'(r'))\cap I(r'))\cup U(r')\right)(r)\\
  &=\left(\lambda m'.\lambda r'.(m'(r')\cap I(r'))\cup U(r')\right)
    \left(\left(\lambda m'.\lambda r'.(m'(r')\cap I'(r'))\cup U'(r')\right)  (m)\right)(r)\\
  &=\left(\denote{\setpair IU}\circ\denote{\setpair{I'}{U'}}\right)(m)(r).\qedhere
\end{align*}
\end{proof}

\equalMeet*
\begin{proof}
  For all $r\in R$ and $m\in R\to2^T$,
\begin{align*}
  \left(\denote{\setpair{I}{U} \sqcap \setpair{I'}{U'}}\right)&(m)(r)\\
  &=(\denote{\setpair{\lambda r\,.\,I(r) \cup I'(r)}{\lambda r\,.\,U(r) \cup U'(r)}})(m)(r)\\
  &=(m(r)\cap(I(r)\cup I'(r))\cup U(r)\cup U'(r)\\
  &=(m(r)\cap I(r))\cup U(r)\cup(m(r)\cap I'(r))\cup U'(r)\\
  &=\denote{\setpair{I}{U}}(m)(r)\cup\denote{\setpair{I'}{U'}}(m)(r)\\
  &=\left(\denote{\setpair{I}{U}} \sqcap \denote{\setpair{I'}{U'}}\right)(m)(r).\qedhere
\end{align*}
\end{proof}

@@

\precision*
\begin{proof}
    \otodo{It's Section 3.2.}
  Let $P$ be an IFDS problem. Recall from Section~\ref{sec:ideToIfds} that the result of an IFDS analysis $\resultIFDS(P)$ maps supergraph
  nodes $n\in N^*$ to sets of data-flow facts $\delta\in2^D$. Specifically,
  \begin{equation}
    \resultIFDS(P)=\lambda n\,.\,\mvp F(n)\,.
  \end{equation}
  \otodo{Where do we mention $\backEq$ in this lemma?}
  The transformation $\backCC$ is the same as $\backEq$, except that it can remove data-flow facts from the result:
  \begin{align*}
    \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
    &=\left(\lambda n'\,.\,\left\{d\,|\,\forall r\in R\,.\,\resultIDE(\transCC_R(P))(n')(d)(r)\ne\top_T\right\}\right)(n)\\
    &=\left\{d\,|\,\forall r\in R\,.\,\resultIDE(\transCC_R(P))(n')(d)(r)\ne\top_T\right\}\\
    &\subseteq\left\{d\,|\,\resultIDE(\transCC_R(P))(n)(d)\ne\topcc\right\}\\
    &=\backEq\left(\resultIDE(\transCC_R(P))(n)\right).\qedhere
  \end{align*}
\end{proof}

@@

\sound*
\begin{proof}
According to~\eqref{eq:ideresult}, we can rewrite~\eqref{eq:transCcDef} as
\begin{align*}\label{eq:transCcResultCc}
  \backCC(\resultIDE(\transCC_R(P)))(n)
  &=\{d'\,|\,\forall r\in R\,.\,\mvp{\textsf{Env}}(n,\,d'))(r)\ne\top_T\}\\
  &=\left\{d'\,|\,\forall r\in R\,.\bigsqcap_{q\in\textsf{VP}(n)}\menv(q)(\Omega)(d')(r)\ne\top_T\right\}\\
  &=\left\{d'\,|\,\forall r\in R\,.\bigsqcap_{q\in\textsf{VP}(n)}\mpddef q {d'}(r)\ne\top_T\right\}.
\end{align*}
  According to~Lemma~\ref{lem:sound3}, since $d\in M_F(p)(\varnothing)$, then for any $r\in R$, $\mpd(r)\ne\top_T$. 
  Since $\mpd(r)$ is a non-empty set that is contained in $\bigsqcap_{q\in\textsf{VP}(n)}\mpddef q {d}(r)$, it follows that $$\bigsqcap_{q\in\textsf{VP}(n)}\mpddef q {d}(r)\ne\top_T.$$ Therefore, $d\in \backCC(\resultIDE(\transCC_R(P)))(n)$.
\end{proof}

### 2014-12-16 content/impl.tex
modifying implementation explanation
repeating lemmas in appendix

To compare two micro-functions for equality, it suffices to compare the corresponding
sets $I(r)$ and $U(r)$ for all receivers $r$. The following lemma shows that this
implementation of comparison corresponds to equality of the represented micro-functions:
\begin{lemma}\label{lem:equalEq}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \forall r\,.\,I(r) = I'(r) \land U(r) = U'(r)
        \iff
        \denote{\setpair{I}{U}} = \denote{\setpair{I'}{U'}}.
    \] 
\end{lemma}
The proofs of all of the lemmas and theorems in the paper are provided in the supplemental
material submitted with the paper.
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}

\otodo{The normalization function $\mathcal N$ was only needed in composition. I therefore
    ``inlined'' it into composition to save space. This makes composition a bit harder to intuitively
    verify, but saves complexity here. The verification can be in the proof in the supplementary appendix.}
The composition of two micro-function representations is defined as follows:
$\setpair{I}{U} \circ \setpair{I'}{U'} = \setpair{\lambda r\,.\,(I(r) \cap I'(r))\cup U(r)}{\lambda r\,.\,(I(r)\cap U'(r))\cup U(r)}$.
The following lemma shows that this implementation corresponds to the composition
of the denoted functions:
\begin{lemma}\label{lem:equalComp}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \denote{\setpair{I}{U} \circ \setpair{I'}{U'}}
        = 
        \denote{\setpair{I}{U}} \circ \denote{\setpair{I'}{U'}}.
    \]
\end{lemma}
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}

The meet of two micro-function representations is defined as follows:
$\setpair{I}{U} \sqcap \setpair{I'}{U'} = \setpair{\lambda r\,.\,I(r) \cup I'(r)}{\lambda r\,.\,U(r) \cup U'(r)}$.
The following lemma shows that this implementation corresponds to the meet
of the denoted functions:
\begin{lemma}\label{lem:equalMeet}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \denote{\setpair{I}{U} \sqcap \setpair{I'}{U'}}
        = 
        \denote{\setpair{I}{U}} \sqcap \denote{\setpair{I'}{U'}},
    \]
    where
    $\denote{\setpair{I}{U}} \sqcap \denote{\setpair{I'}{U'}}
     =\lambda m.\lambda r.\denote{\setpair{I}{U}}(m)(r)\cup\denote{\setpair{I'}{U'}}(m)(r)$.\mtodo{I added the definition of the meet of two functions.}
\end{lemma}
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}

@@

\begin{lemma}[Precision]\label{lem:subsetifds}
    For an IFDS problem $P$ and all ${n\in N^*}$,
    \begin{equation}\label{eq:correct}
      \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
      \subseteq
      \resultIFDS(P)(n)\,.
    \end{equation}
\end{lemma}

@@

\begin{lemma}[Soundness]\label{lem:sound}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  \begin{equation}
    d\in \backCC\left(\resultIDE(\transCC_R(P))\right)(n)\,.
  \end{equation}
\end{lemma}

@@

\subsection{Implementation of the Correlated-Calls Analysis}
The correlated-calls analysis was implemented in the Scala programming language~\cite{odersky2004overview}. 
We chose Java as the target language for client programs of the analysis.
To retrieve information about an input program, such as its control-flow supergraph or the set of receivers and their types, we used the WALA framework for static analysis on Java bytecode~\cite{fink2012wala}.

Since WALA currently only contains an implementation of IFDS, we implemented IDE from scratch. Instead of using WALA's IFDS implementation, to run an IFDS problem, we converted it to an IDE problem and used our own IDE solver.

@@

>>>>

To compare two micro-functions for equality, it suffices to compare the corresponding
sets $I(r)$ and $U(r)$ for all receivers $r$. The following lemma shows that this
implementation of comparison corresponds to equality of the represented micro-functions:
\begin{restatable}{lemma}{equalEq}
\label{lem:equalEq}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \forall r\,.\,I(r) = I'(r) \land U(r) = U'(r)
        \iff
        \denote{\setpair{I}{U}} = \denote{\setpair{I'}{U'}}.
    \] 
\end{restatable}
The proofs of all of the lemmas and theorems in the paper are provided in the supplemental
material submitted with the paper.
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}

\otodo{The normalization function $\mathcal N$ was only needed in composition. I therefore
    ``inlined'' it into composition to save space. This makes composition a bit harder to intuitively
    verify, but saves complexity here. The verification can be in the proof in the supplementary appendix.}
The composition of two micro-function representations is defined as follows:
$\setpair{I}{U} \circ \setpair{I'}{U'} = \setpair{\lambda r\,.\,(I(r) \cap I'(r))\cup U(r)}{\lambda r\,.\,(I(r)\cap U'(r))\cup U(r)}$.
The following lemma shows that this implementation corresponds to the composition
of the denoted functions:
\begin{restatable}{lemma}{equalComp}\label{lem:equalComp}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \denote{\setpair{I}{U} \circ \setpair{I'}{U'}}
        = 
        \denote{\setpair{I}{U}} \circ \denote{\setpair{I'}{U'}}.
    \]
\end{restatable}
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}

The meet of two micro-function representations is defined as follows:
$\setpair{I}{U} \sqcap \setpair{I'}{U'} = \setpair{\lambda r\,.\,I(r) \cup I'(r)}{\lambda r\,.\,U(r) \cup U'(r)}$.
The following lemma shows that this implementation corresponds to the meet
of the denoted functions:
\begin{restatable}{lemma}{equalMeet}\label{lem:equalMeet}
\label{lem:equalMeet}
    For any pair of micro-function representations $\setpair{I}{U}$, $\setpair{I'}{U'}$,
    \[
        \denote{\setpair{I}{U} \sqcap \setpair{I'}{U'}}
        = 
        \denote{\setpair{I}{U}} \sqcap \denote{\setpair{I'}{U'}},
    \]
    where
    $\denote{\setpair{I}{U}} \sqcap \denote{\setpair{I'}{U'}}
     =\lambda m.\lambda r.\denote{\setpair{I}{U}}(m)(r)\cup\denote{\setpair{I'}{U'}}(m)(r)$.\mtodo{I added the definition of the meet of two functions.}
\end{restatable}
\otodo{Prove this lemma in the appendix. The parts of the proof can be found in section 5.}

@@

\begin{restatable}[Precision]{lemma}{precision}\label{lem:subsetifds}
    For an IFDS problem $P$ and all ${n\in N^*}$,
    \begin{equation}\label{eq:correct}
      \backCC\left(\resultIDE(\transCC_R(P))\right)(n)
      \subseteq
      \resultIFDS(P)(n)\,.
    \end{equation}
\end{restatable}

@@

\begin{restatable}[Soundness]{lemma}{sound}\label{lem:sound}
  Let $p=[\startmain,\,\dots,\,n]$ be a concrete execution path, and let $d\in D$.
  If $d\in M_F(p)(\varnothing),$
  then
  \begin{equation}
    d\in \backCC\left(\resultIDE(\transCC_R(P))\right)(n)\,.
  \end{equation}
\end{restatable}

@@

\subsection{Implementation of the Correlated-Calls Analysis}
The correlated-calls analysis was implemented in the Scala programming language~\cite{odersky2004overview}. 
We chose Java as the target language for client programs of the analysis.
To retrieve information about an input program, such as its control-flow supergraph or the set of receivers and their types, we used the WALA framework for static analysis on Java bytecode~\cite{fink2012wala}.
Since WALA currently only contains an implementation of IFDS, we implemented IDE from scratch. Instead of using WALA's IFDS implementation, to run an IFDS problem, we converted it to an IDE problem and used our own IDE solver.

@@

The artifact of our implementation can be used to solve any IFDS problem. In the presence of correlated calls, the result of the correlated-calls analysis can be more precise than the result obtained with an IFDS solver.
